Hydras \& Co.

### 2022-05-09 doc/JFLA2022/Slides/Slides.tex
corrections in tex files

  %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}
  \begin{figure}[ht]
\centering
\scalebox{0.7}{
\begin{tikzpicture}[thick]

  \begin{scope}[yshift=2.0cm, xshift=1.7cm] 
     \node[ellipse, dotted, fill=orange!20, minimum height=0.5cm,minimum width=1.5cm] (settheory) { Set theory\footnote{N. Bourbaki}};
     \node[ellipse, dotted, fill=orange!20, minimum height=0.5cm,minimum width=1.5cm, left of=settheory, node distance=8.4cm] (prooftheory) { Proof theory\footnote{K. Schutte, J. Ketonen, R. Solovay, L. Kirby \& J. Paris}};
  \node[ellipse, dotted, fill=orange!20, minimum height=0.5cm,minimum width=1.5cm, left of=settheory, node distance=4.8cm] (algo) { Algorithmics\footnote{S. Brlek \emph{et. al}}};
\end{scope}

\begin{scope}[yshift=1.0cm, xshift=-1.5cm]
   \node[rectangle, dotted, draw= black!20,fill=green!20,minimum height=0.5cm,minimum width=1.5cm] (acl2) { [ACL2]\footnote{P. Manolios \&  D. Vroon}};   
\end{scope}

\begin{scope}[xshift=-6.1cm]
\node[rectangle, dotted,draw=black,minimum height=0.5cm,minimum width=1.5cm] (goedel) { Goedel\footnote{R. O'Connor} };
\node[rectangle, dotted,draw=black,minimum height=0.5cm,minimum width=1.5cm,left of=goedel, node distance=2.2cm] (pock) { Pocklington };
\node[rectangle, dotted,draw=black,minimum height=0.5cm,minimum width=1.5cm,right of=goedel, node distance=2.5cm] (additions) { Additions };
\node[rectangle, dotted,draw=black,minimum height=0.5cm,minimum width=1.5cm,right of=additions, node distance=2cm] (cantor) { Cantor\footnote{C. \& E. Contejean} };

\node[rectangle, dotted,draw=black,minimum height=0.5cm,minimum width=1.5cm,right of=cantor, node distance=3.2cm] (cats) { Categories in ZFC };
\end{scope}

\begin{scope}[yshift=-1cm,xshift=0.45cm]
  \node[rectangle, draw=black,fill=blue!30,minimum height=0.5cm,minimum width=1.5cm] (paramcoq) { Paramcoq };
\node[rectangle, draw=black,minimum height=0.5cm,minimum width=1.5cm,right of=paramcoq, node distance=2.5cm] (mathcomp) { MathComp };
\node[rectangle, draw=black,minimum height=0.5cm,minimum width=1.5cm,node distance=2cm,left of=paramcoq, node distance=4.1cm] (equations) { Equations };
\end{scope}

\begin{scope}[yshift=-2cm,xshift=-4cm]
\node[rectangle, draw=black,fill=orange!80,minimum height=0.5cm,minimum width=1.5cm] (hydras) { Hydra-battles };
\node[rectangle, draw=black,fill=blue!30,minimum height=0.5cm,minimum width=1.5cm,left of=hydras, node distance=3.5cm] (pockcc) { Pocklington };
\node[rectangle, draw=black,fill=orange!80,minimum height=0.5cm,minimum width=1.5cm,right of=hydras, node distance=2.5cm] (chains) { Addition-chains };
\node[rectangle, draw=black,fill=blue!30,minimum height=0.5cm,minimum width=1.5cm,right of=chains, node distance=2.2cm] (gaia) { Gaia\footnote{J. Grimm} };
\end{scope}

\begin{scope}[yshift=-3cm,xshift=-6cm]
 \node[rectangle, draw=black,fill=blue!30,minimum height=0.5cm,minimum width=1.5cm] (goedelcc) { Goedel };
 \node[rectangle, draw=black,fill=orange!80,minimum height=0.5cm,minimum width=1.5cm,right of=goedelcc, node distance=3cm] (gaiahydras) { Gaia-hydras };
\end{scope}

\draw[lookcolor,->,dotted] (acl2.south) -- (cantor.north) ;
\draw[lookcolor,->,dashed] (additions.south) -- (chains.north) ;
\draw[lookcolor,->,dashed] (goedel) -- (goedelcc) ;
\draw[lookcolor,->,dashed] (goedel) -- (hydras.north west) ;
\draw[lookcolor,->,dashed] (cantor) -- (hydras.north east) ;
\draw[lookcolor,->,dashed] (cantor.south) -- (gaia.north west) ;
\draw[lookcolor,->,dashed] (pock) -- (pockcc) ;
\draw[lookcolor,->,dashed] ([xshift=2.1cm]cats) -- (gaia) ;

\draw[->] (hydras) -- (gaiahydras) ;
\draw[->] (hydras.south west) -- (goedelcc) ;
\draw[->] (gaia.south west) -- (gaiahydras.east) ;

\draw[->] (equations) -- (hydras) ;
\draw[->] (mathcomp) -- (gaia) ;
\draw[->] (mathcomp) -- (chains) ;
\draw[->] (paramcoq.south) -- (chains) ;
\draw[->] (pockcc) -- (goedelcc) ;
\draw[dotted, ->] (prooftheory) -- (hydras);
\draw[dotted, ->] (settheory) -- (cats);
\draw[dotted, ->] (prooftheory) -- (goedel);
\draw[dotted, ->] (algo) -- (additions);
\end{tikzpicture}
}
\caption{Genealogy and dependencies for \Hydras packages.}
  \label{fig:genealogy}
\end{figure}
\end{frame}

@@

\begin{frame}
   \begin{block}{An example-driven book}
    \begin{itemize}
    \item The structure is determined by the presented topics (discrete math, algorithmics, etc.). \coq technology (formalization and proof patterns) is presented when needed.
    \item No need to be complete (there is already a reference manual)
 %   \item No need to introduce \coq (there are already a lot of initiation books and tutorials)
    \item Comparison of  different formalizations of a given notion, and several proofs of the same theorem.
      \item Show also interesting dead ends (formalization, tactics, proof structure).
    \end{itemize}
  \end{block}
  % \begin{block}{Examples}
  %   \begin{itemize}
  %   \item Tsukuba Coq Users's Group: regular languages, $\lambda$-calculus, arithmetic, ... (in Japanese + ssreflect)
  %   \item Software Foundations (B. Pierce \emph{et al.})
  %   \end{itemize}
  % \end{block}
\end{frame}

@@

\begin{frame}
  \frametitle{What hydras are talking about?}
  \begin{block}{}
    \Hydras presents a consistent set of examples, which allow us to formalize 
    some \textcolor{mathcolor}{discrete math}, discuss
    \textcolor{coqstylecolor}{\coq style, specification and proof patterns}, and depend on \textcolor{plugincolor}{a few plug-ins}.
  \end{block}
  \begin{block}{}
    \begin{description}
    \item[hydra battles:]

@@

\begin{frame}
  \frametitle{Example: Kirby \& Paris' hydras}
  \begin{figure}[h]
    \centering
    \begin{tikzpicture}[very thick, scale=0.25]
      \node (h1) at (4,0){$\bullet$};
      \node[blue] (h2) at (4,2){$\bullet$};
      \node[blue] (h3) at (2,4){$\bullet$};
      \node[blue] (h4) at (0,6){$\bullet$};
      \node[red!90!black] (h7) at (6,2){$\bullet$};
      \node[blue] (h5) at (4,6){$\bullet$};
      \node[blue] (h6) at (6,4){$\bullet$};
      \draw (h1) -- (h2) ;
      \draw (h2) -- (h3) ;
      \draw (h1) -- node[red!90!black,font=\small,sloped,
      shift={(0.01,-0.075)},rotate=90]{\textbf{\ding{34}}} (h7);
      \draw (h3) -- (h4) ;
      \draw (h3) -- (h5) ;
      \draw (h2) -- (h6);

@@

\begin{frame}
  \begin{block}{}
    Using  \texttt{PrimRec\_PrimRecs\_ind}, we prove the following lemma:

Comparison of the length of hydra-battles (w.r.t to the initial replication factor) with the diagonalized Ackermann function proves the theorem.
  \end{block}
\end{frame}

>>>>

\end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}
  \frametitle{What hydras are talking about?}
  \begin{block}{}
    \Hydras presents a consistent set of examples, which allow us to formalize 
    some \textcolor{mathcolor}{discrete math}, discuss
    \textcolor{coqstylecolor}{\coq style, specification and proof patterns}, and depend on \textcolor{plugincolor}{a few plug-ins}.
  \end{block}
  \begin{block}{}
    \begin{description}
    \item[hydra battles:]

  %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}
\centering
\scalebox{0.7}{
\begin{tikzpicture}[thick]

  %%%%%%%%%%%% First line %%%%%%%%%%%%%%%%
  \begin{scope}[yshift=2.0cm, xshift=3cm]
     \node[ellipse, dotted, fill=orange!20, minimum height=0.5cm,minimum width=1.5cm] (algo) { Algorithmics};
     \node[ellipse, dotted, fill=orange!20, minimum height=0.5cm,minimum width=1.5cm, left of=algo, node distance=6cm] (settheory) { Set theory};
     \node[ellipse, dotted, fill=orange!20, minimum height=0.5cm,minimum width=1.5cm, left of=settheory, node distance=6cm] (prooftheory) {Proof theory};

\end{scope}

%%%%%% ACL2   %%%%%%%

\begin{scope}[yshift=1.0cm, xshift=-3.7cm]
   \node[rectangle, dotted, draw= black!20,fill=green!20,minimum height=0.5cm,minimum width=1.5cm] (acl2) { [ACL2]\footnote{P. Manolios \&  D. Vroon}};   
\end{scope}

%%%%%%% Contribs %%%%%%%%%%%%%%%%%%%%

\begin{scope}[xshift=-5.7cm]
\node[rectangle, dotted,draw=black,minimum height=0.5cm,minimum width=1.5cm] (goedel) { Goedel\footnote{R. O'Connor} };
\node[rectangle, dotted,draw=black,minimum height=0.5cm,minimum width=1.5cm,left of=goedel, node distance=2.2cm] (pock) { Pocklington };
\node[rectangle, dotted,draw=black,minimum height=0.5cm,minimum width=1.5cm,right of=goedel, node distance=2.5cm] (cantor) { Cantor\footnote{C. \& E. Contejean}};

\node[rectangle, dotted,draw=black,minimum height=0.5cm,minimum width=1.5cm,right of=cantor, node distance=3.6cm] (cats) { Categories in ZFC };
\node[rectangle, dotted,draw=black,minimum height=0.5cm,minimum width=1.5cm,right of=cats, node distance=3.2cm] (additions) {Additions};

\end{scope}

%%%%%%%%%%%%%%%%% Plug-ins %%%%%%%%%%%%%%

\begin{scope}[yshift=-0.8cm,xshift=-1.45cm]
\node[rectangle, draw=black,minimum height=0.5cm,minimum width=1.5cm] (mathcomp) { MathComp };

  \node[rectangle, draw=black,fill=blue!30,minimum height=0.5cm,minimum width=1.5cm, right of=mathcomp, node distance=5.4cm] (paramcoq) { Paramcoq};

\node[rectangle, draw=black,minimum height=0.5cm,minimum width=1.5cm,node distance=2cm,left of=mathcomp, node distance=7.7cm] (equations) { Equations };
\end{scope}

\begin{scope}[yshift=-2cm,xshift=-4.2cm]
\node[rectangle, draw=black,fill=orange!80,minimum height=0.5cm,minimum width=1.5cm] (hydras) { Hydra-battles };
\node[rectangle, draw=black,fill=blue!30,minimum height=0.5cm,minimum width=1.5cm,left of=hydras, node distance=3.5cm] (pockcc) { Pocklington };
\node[rectangle, draw=black,fill=blue!30,minimum height=0.5cm,minimum width=1.5cm, right of=hydras, node distance=3.5cm] (gaia) { Gaia\footnote{J. Grimm} };

\node[rectangle, draw=black,fill=orange!80,minimum height=0.5cm,minimum width=1.5cm,right of=gaia, node distance=3.5cm] (chains) { Addition-chains };

\end{scope}

%%%%%%%%%%%%%%%%% goedel, etc . %%%%%%%%%%%%%%%
\begin{scope}[yshift=-3.7cm,xshift=-6cm]
 \node[rectangle, draw=black,fill=blue!30,minimum height=0.5cm,minimum width=1.5cm] (goedelcc) { Goedel };
 \node[rectangle, draw=black,fill=orange!80,minimum height=0.5cm,minimum width=1.5cm,right of=goedelcc, node distance=3cm] (gaiahydras) { Gaia-hydras };
\node[rectangle, draw=black,minimum height=0.5cm,minimum width=1.5cm,right of=gaiahydras, node distance=5.1cm] (alectryon) { Alectryon};

\end{scope}

\begin{scope}[yshift=-5cm,xshift=4cm]
\node[rectangle, draw=black,fill=red!80,minimum height=0.5cm,minimum width=1.5cm] (book) { \Hydras};  
\end{scope}

\draw[lookcolor,->,dotted] (acl2.south) -- (cantor.north) ;
\draw[lookcolor,->,dashed] (additions.south) -- (chains.north) ;
\draw[lookcolor,->,dashed] (goedel) -- (goedelcc) ;
\draw[lookcolor,->,dashed] (goedel) -- (hydras.north) ;
\draw[lookcolor,->,dashed] (cantor) -- (hydras.north east) ;
\draw[lookcolor,->,dashed] (cantor.south) -- (gaia.north west) ;
\draw[lookcolor,->,dashed] (pock) -- (pockcc) ;
\draw[lookcolor,->,dashed] ([xshift=2.1cm]cats) -- (gaia) ;

\draw[->] (hydras.south) -- (gaiahydras) ;
\draw[->] (hydras.south west) -- (goedelcc) ;
\draw[->] (gaia.south west) -- (gaiahydras.north) ;

\draw[->] (equations) -- (hydras) ;
\draw[->] (mathcomp) -- (gaia) ;
\draw[->] (mathcomp) -- (chains) ;
\draw[->] (paramcoq.south) -- (chains) ;
\draw[->] (pockcc) -- (goedelcc) ;
\draw[dotted, ->] (prooftheory) -- (hydras);
\draw[dotted, ->] (settheory) -- (cats);
\draw[dotted, ->] (prooftheory) -- (goedel);
\draw[dotted, ->] (algo) -- (additions);
\draw[->, red] (alectryon) -- (book) ;
\draw[->, red] (chains) -- (book) ;
\draw[->, red] (gaiahydras) -- (book) ;
\draw[->, red] (hydras.south) -- (book) ;

\end{tikzpicture}
}
% \caption{Genealogy and dependencies for \Hydras.}
%   \label{fig:genealogy}
% \end{figure}
\end{frame}

@@

% \begin{frame}
%    \begin{block}{An example-driven book}
%     \begin{itemize}
%     \item The structure is determined by the presented topics (discrete math, algorithmics, etc.). \coq technology (formalization and proof patterns) is presented when needed.
%     \item No need to be complete (there is already a reference manual)
%  %   \item No need to introduce \coq (there are already a lot of initiation books and tutorials)
%     \item Comparison of  different formalizations of a given notion, and several proofs of the same theorem.
%       \item Show also interesting dead ends (formalization, tactics, proof structure).
%     \end{itemize}
%   \end{block}
%   % \begin{block}{Examples}
%   %   \begin{itemize}
%   %   \item Tsukuba Coq Users's Group: regular languages, $\lambda$-calculus, arithmetic, ... (in Japanese + ssreflect)
%   %   \item Software Foundations (B. Pierce \emph{et al.})
%   %   \end{itemize}
%   % \end{block}
% \end{frame}

@@

@@

\begin{frame}
  \frametitle{Kirby \& Paris' hydras}
  \begin{figure}[h]
    \centering
    \begin{tikzpicture}[very thick, scale=0.25]
      \node (h1) at (4,0){$\bullet$};
      \node[blue] (h2) at (4,2){$\bullet$};
      \node[blue] (h3) at (2,4){$\bullet$};
      \node[blue] (h4) at (0,6){$\bullet$};
      \node[red!90!black] (h7) at (6,2){$\bullet$};
      \node[blue] (h5) at (4,6){$\bullet$};
      \node[blue] (h6) at (6,4){$\bullet$};
      \draw (h1) -- (h2) ;
      \draw (h2) -- (h3) ;
      \draw (h1) -- node[red!90!black,font=\small,sloped,
      shift={(0.01,-0.075)},rotate=90]{\textbf{\ding{34}}} (h7);
      \draw (h3) -- (h4) ;
      \draw (h3) -- (h5) ;
      \draw (h2) -- (h6);

@@

### 2022-05-09 doc/gaia-chapter.tex
corrections in tex files

%\inputsnippets{T1Bridge/utf8try}

@@

The definition of the binary operations $+$ and $\times$
requires the comparison of ordinal terms, which is not implemented the same way in both libraries. Thus, the proof of the following lemmas applies applies
compatibility lemmas like \texttt{compare\_ref} (see Section~\vref{sect:gaia-compare-ref}).

@@

\subsubsection{Remark}
\label{sect:gaia-masking}
We try to respect the following ``protocol''.
Let us assume we adapt some definitions and properties
from a given module \texttt{hydras.Epsilon0.Foo}, and a constant \texttt{bar} of this module, whose absolute name is
thus \texttt{hydras.Epsilon0.Foo.bar}.
\begin{enumerate}
\item We create a module \texttt{gaia\_hydras.GFoo}
  (``G'' for \gaia).
\item Inside \texttt{GFoo}, the identifier \texttt{bar}
  is reserved in priority to \gaia's version.
\HydrasLib' constructions are available through qualified names, like \texttt{Epsilon0.Foo.bar}, \texttt{Foo.bar}, or even a short notation like \texttt{hbar}  (``h'' for \HydrasLib).
\end{enumerate}

>>>>

@@

The definition of the binary operations $+$ and $\times$
requires the comparison of ordinal terms, which is not implemented in \gaiaHydras  the same way as in \HydrasLib. Thus, the proof of the following lemmas applies applies
compatibility lemmas like \texttt{compare\_ref} (see Section~\vref{sect:gaia-compare-ref}).

@@

\subsubsection{Remark}
\label{sect:gaia-masking}
We try to respect the following convention:
Let us consider  some module \texttt{hydras.Epsilon0.Foo}.
Its adaptation to \gaia will be a new module called \linebreak
\texttt{gaia\_hydras.GFoo} (``G'' for \gaia).

Let \texttt{bar} be a symbol defined in \texttt{Foo}.
Inside \texttt{GFoo}, the identifier \texttt{bar}
  is reserved in priority to \gaia's adaptation.
  \HydrasLib' definition is still available through qualified names, like \texttt{Epsilon0.Foo.bar}, \texttt{Foo.bar}, or even a short notation like \texttt{hbar}  (``h'' for \HydrasLib).
